---
title: "p8105_hw1_hc3654"
author: "Haohua Chen"
date: "2025-09-25"
output: github_document
  
---

```{r}
library(tidyverse)
library(readxl)
library(haven)
```


# Problem 1


## pols-month dataset


```{r}
pols_month_df <- 
  read_csv("fivethirtyeight_datasets/pols-month.csv", na = c("NA", ".", " ")) |> 
  janitor::clean_names() |> 
  separate(mon , into = c("year", "month", "day"), sep = "-"
  ) |> 
  mutate(month = month.name[as.numeric(month)]) |> 
  mutate(president = case_when(
    prez_dem == 1 ~ "dem",
    prez_gop == 1 ~ "gop"
  )) |> 
  select(-prez_dem, -prez_gop, -day)

  
```

## snp dataset


```{r}
snp_df <- 
  read_csv("fivethirtyeight_datasets/cpi.csv") |> 
  janitor::clean_names() |> 
  separate(date, into = c("year", "month", "day"), sep = "-") |> 
  mutate(month = month.name[as.numeric(month)]) |> 
  rename(snp_value = value) |> 
  arrange(year,month)

```

## unemployment dataset


```{r}
unemp_df <- 
  read_csv("fivethirtyeight_datasets/unemployment.csv") |> 
  janitor::clean_names() |> 
  pivot_longer(
    cols = jan:dec,
    names_to = "month",
    values_to = "unemployment"
  ) |> 
  mutate(
    month = month.name[match(tolower(month), tolower(month.abb))]) |> 
  mutate(year = as.character(year))

```
## Merge


```{r}
pols_snp_unemp <- 
  inner_join(pols_month_df,snp_df, by = c("year","month")) |> 
  inner_join(unemp_df, by = c("year", "month"))

knitr::kable(head(pols_snp_unemp,10), caption  = "First 10 rows")

```


## Writing question for problem 1:

The pols-month.csv dataset contains 822 observations and 9 variables. It tracks the counts of Republican and Democratic governors, senators, and representatives (gov_gop, sen_gop, rep_gop, gov_dem, sen_dem, rep_dem). It also shows which party the sitting president belongs to, with year and month as key time variables.

The snp.csv dataset contains 822 observations and 4 variables. It represents the monthly S&P stock market index values (snp_value), along with date variables (year, month, day).

The unemployment.csv dataset contains 816 observations and 3 variables, recording the U.S. monthly unemployment rate (unemployment) by year and month.

After merging the three datasets, we can easily track unemployment rates and economic changes in relation to political changes.





# Problem 2


## Setting up datasets


```{r}
MrTrashWheel_df <- 
  read_excel("202509 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel" , na = c("NA", ".", " ")) |> 
  janitor::clean_names() |> 
  filter( dumpster != "na") |> 
  select(-15,-16) |> 
  mutate(sports_balls = as.integer(round(sports_balls))) |> 
  mutate(track = "Mr.Trash Wheel") |> 
  mutate(year = as.numeric(year))

ProfTrashWheel_df <- 
  read_excel("202509 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", na = c("NA", ".", " ")) |> 
  janitor::clean_names() |> 
  filter( dumpster != "na") |> 
  mutate(track = "Professor Trash Wheel")


GwynndaTrashWheel_df <- 
  read_excel("202509 Trash Wheel Collection Data.xlsx", sheet = "Gwynns Falls Trash Wheel", na = c("NA", ".", " ")) |> 
janitor::clean_names()|> 
  filter( dumpster != "na") |> 
  mutate(track = "Gwynnda Trash Wheel")
```


## Combines datasets


```{r}
AllTrashWheel_df <- 
  bind_rows(MrTrashWheel_df,ProfTrashWheel_df,GwynndaTrashWheel_df)

knitr::kable(head(AllTrashWheel_df,10), caption = "First 10 rows")
```

# Writing question for Problem 2:

The total weight of the Professor Trash Wheel is `r sum(ProfTrashWheel_df$weight_tons, na.rm = TRUE)` tons during those days. And 
And the total number of cigarette butts collected by Gwynnda in June of 2022 is  `r GwynndaTrashWheel_df |> 
    dplyr::filter(year == 2022, month == "June") |> 
    dplyr::summarise(total = sum(cigarette_butts, na.rm = TRUE)) |> 
    dplyr::pull(total) |> 
    format(scientific = FALSE, big.mark = ",")`.


# Problem 3


## Clean data

```{r}
Zip_Codes_df <- 
  read_csv("zillow_data/Zip Codes.csv", na = c("NA", ".", " ")) |> 
  janitor::clean_names() |> 
  arrange(zip_code)

Zillow_df <- 
  read_csv("zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv", na = c("NA", ".", " ")) |> 
  janitor::clean_names() |> 
  pivot_longer(
    cols = x2015_01_31 : x2024_08_31,
    names_to = "Date",
    values_to = "rental_fee"
  ) |> 
  separate(Date, into = c("year", "month" , "day"), sep = "_") |> 
  mutate(year = sub("x","",year)) |> 
  mutate(county = str_replace(county_name," County","")) |> 
  select(-county_name) |> 
  arrange(desc(rental_fee)) |> 
  rename( zip_code = region_name) 
```

## Create single 

```{r}
Zillow_Zipcode_df <-  
  left_join(Zip_Codes_df, Zillow_df, by = c("zip_code","county")) |> 
  select(-county_fips, -state_fips,-county_code,-file_date, -region_id, -size_rank, -region_type, -metro, -state_name)


  
  
  
knitr::kable(head(Zillow_Zipcode_df,10), caption = "First 10 rows")
```

## Writing question for Problem 3: 
```{r}
distinct_all <- Zillow_Zipcode_df %>% distinct(zip_code)
distinct_Zillow <- Zillow_df %>% distinct(zip_code)

Exclude_Zillow <- setdiff(Zip_Codes_df$zip_code, Zillow_df$zip_code)
  
```

```{r}
price_diff <- Zillow_Zipcode_df |> 
  select(year,month, zip_code, neighborhood, rental_fee, county) |> 
  filter(year == '2020' & month == "01" | year == '2021' & month == "01") |> 
  mutate(rental_2020 = case_when(year == '2020' ~ rental_fee), rental_2021 = case_when(year == "2021" ~ rental_fee)) |> 
   pivot_wider(
    id_cols   = c(zip_code, neighborhood, county),   
    names_from  = year,                                  
    values_from = rental_fee,                           
    names_prefix = "rental_",                             
    values_fn  = list(rental_fee = ~ first(na.omit(.))),  
    values_fill = NA_real_                                
  ) |> 
    mutate(diff = rental_2020 - rental_2021) |> 
  arrange(desc(diff))

top10 <- price_diff |> 
  slice_max(order_by = diff, n = 10) |> 
  rename(
    borough = county,
    zip = zip_code,
    neighborhood = neighborhood,
    rent_2020 = rental_2020,
    rent_2021 = rental_2021,
    drop = diff
  )

```




There are total **`r nrow(Zillow_Zipcode_df)` obs, `r dplyr::n_distinct(Zillow_Zipcode_df$zip_code)`** unique zip codes and **`r dplyr::n_distinct(Zillow_Zipcode_df$neighborhood)`** unique neighborhoods（include an "na").
ZIP codes appear in the ZIP code dataset but not in the Zillow Rental Price dataset are: `r paste(Exclude_Zillow, collapse = " ")`.

**Some Example should be consider to this results can be:**

**1.** Some of the Zip-code area such as 10020 in NYC are mostly Commercial place.

**2.** Some Zip-code are not for the residential area, For example 10101 - 10199 are mostly using for P.O. Boxes or other institution.

**3.** Some Zip-Code area are very small.

**4.** The data from Zillow is basicly rely on the house for sale or rent.

**The table includes the ten zip codes with largest drop in price is:** `r knitr::kable(top10, digits = 2)`.




